### 2011-11-06 doc/pldi2012/circuit.tex

\section{Example: Circuit Simulator}

@@

Listing~\ref{lst:code_ex} shows an example of Legion code for an electrical
circuit simulation.  Although it does not use all the capabilities of Legion
(or all the aspects of the language), it covers the core set.  Legion code is
derived from C, but different in several important ways.

Lines 1-16 
define the data structures used in the rest of the code.  A {\tt Circuit}
is an irregular data structure that describes the wires in the circuit and
the 'nodes' (where the wires meet).  The basic nodes and wires are defined
first.  Nodes have a few fields for the simulation (e.g. charge, capacitance),
and are hooked up in a circular linked list to allow the simulation to be
able to iterate over all of them.  Pointers in data structures (e.g. the
``next'' fields in this linked list) must be constrained
to a particular region, so Legion uses a {\tt Type@region} syntax rather than
C's unconstrained {\tt Type~*}.  The example relies on the ability to
constrain different linked lists of nodes differently, so rather than using
the name of some static region the {\tt Node} definition is parameterized on
a region {\tt rn}.  The use of the same parameter on the type pointed to by
the next pointer means that every single node in the linked list is constrained
to reside in the specified region.

The definition of a {\tt Wire} is similar, except that a wire must refer to
the two nodes to which it is electrically connected.  The pointers to the nodes
must be constrained, and the type that the pointer points to requires a region
as well, but unlike in the linked list case, two different regions are used.
This is a common occurrance in Legion code, and arises when one needs to place
one constraint on the objects directly referenced, and a second (presumably
weaker) constraint on objects that are indirectly referencable through the
direct pointers.

With the definition of nodes and wires out of the way, lines 4-9 declare how
they are formed into a {\tt Circuit}, and is our first example of a
\emph{region relationship}.  At the language level, a region relationship is
similar to a structure, except that it is allowed to contain references to
dynamically allocated regions as fields, and the types of the fields of a region
relationship are allowed to refer to those region fields.  (This will be 
discussed in additional detail below.)  It is also allowed
to place constraints on those region fields, which we'll see in the next
declaration.  In this example, a circuit is defined
by two regions, one for its nodes and one for its wires.  The element type of
the node region (and of the {\tt first\_node} pointer) guarantees that every
node in the circuit reside in this region.  Simiarly, the type of the wire
region's element (and of the {\tt first\_wire} pointer) not only guarantees
that every wire in the circuit is in the wire region, but also guarantees that
every node referred to by any wire is in the circuit's node region.

The last definition is for the pieces that the circuit is partitioned into.
The original node and wire regions are specified as parameters, and the
four kinds of subregions that are created below are made fields of the
{\tt CircuitPiece}.  Constraints are placed on all four of these fields that
they must be subregions of the original node and wire regions that were
provided as parameters.  Since the simulation code will need to be able to
iterate over the nodes owned by a piece, the parameterization of the {\tt rn\_pvt},
{\tt rn\_shr}, and {\tt first\_node} fields constrains the linked list of
nodes to reside entirely in those two regions.  The linked list of wires is
similarly constrained, and the wire's node pointers are constrained to fall
into one of the three subregions defined in the {\tt CircuitPiece}.  Finally,
the use of {\tt rn} for {\tt rn\_ghost}'s parameter and the second parameter
in the wire type places no constraints on which nodes a ghost node might point
to.

>>>>

\section{Example: Circuit Simulator}
\label{sec:ex}

@@

In this section we give an informal overview of Legion through an example.
Listing~\ref{lst:code_ex} shows Legion code for an electrical
circuit simulation, which illustrates the core features of the programming model.

Lines 1-16 define data structures.  A circuit consists of wires and
nodes where wires meet.  {\tt Node}s have fields for the simulation
(e.g. charge, capacitance) and are linked in a circular list.
Pointers in data structures (e.g. the {\tt next} fields in the list)
point to a particular region given by the standard {\tt Type@region}
syntax.  The program relies on different lists being in different
regions, so the {\tt Node} definition is parameterized on a region
{\tt rn}.  The {\tt next} field of a $\tt Node \langle rn \rangle$ has
type $\tt Node \langle rn \rangle \mbox{\tt @} rn$, which implies that
all nodes in the list are in region {\tt rn}.

{\tt Wire}s are similar, except that a wire refers to two nodes to
which it is electrically connected.  The two {\tt Node}s are in region {\tt rn},
but they may point to {\tt Node}s in a different region {\tt rn2}.  This is a common pattern in Legion code, and
arises when one needs to place one constraint on the objects directly
referenced, and a second (presumably weaker) constraint on objects
that are indirectly reached through the direct pointers.

Lines 4-9 declare {\tt Circuit}s, which is our first example of a
\emph{region relationship}.  A region relationship is just a structure
with named fields, except that any region fields can be used in the
type declarations.  Thus the region {\tt r\_all\_nodes} contains {\tt
Node}s that point only into region {\tt r\_all\_nodes}, guaranteeing that
all {\tt Node}s are in region {\tt r\_all\_nodes}.  (Note that regions are
declared with the single data type they can contain.)  Simiarly, the
type of the wire region's element not only guarantees that every wire
in the circuit is in the wire region, but also guarantees that every
node referred to by any wire is in the circuit's node region.

The last definition is for the pieces into which the circuit is partitioned.
The original node and wire regions are specified as parameters, and the
four kinds of subregions that are created are made fields of the
{\tt CircuitPiece}.  This declaration also illustrates {\em subregion constraints}
placed on the three region fields stating the region fields  must be subregions of the 
node and wire regions provided as parameters.  Since the simulation code will need to be able to
iterate over the nodes owned by a {\tt CircuitPiece}, the parameterization of the {\tt rn\_pvt} (private nodes),
{\tt rn\_shr} (shared nodes), and {\tt first\_node} fields constrains the linked list of
nodes to reside entirely in those two regions.  The linked list of wires is
similarly constrained, and the wire's node pointers are constrained to fall
into one of the three subregions defined in the {\tt CircuitPiece}.  Finally,
the use of {\tt rn} for {\tt rn\_ghost}'s parameter and the second parameter
in the wire type places no constraints on which nodes a ghost node might point
to.
